Extended Data

### 2021-11-16 doc/manuscript/Dow_Extended_Data.Rmd
reviewed Extended Data Figures 4-5

![**Extended Data Figure 2 | Phenological patterns of forest canopy greenness (top row) and stem growth of ring- and diffuse-porous trees, represented as both relative and cumulative fractions of total annual growth (middle and bottom rows, respectively), at the Smoithsonian Conservation Biology Institute (SCBI) and Harvard Forest.** In the top row, canopy greenness is characterized using the two band Enhanced Vegetation Index (EVI2), with each line representing a year between 2000 and 2018. For stem growth, each line represents the growth of one tree over one year, as modeled based on a five-parameter logistic growth model to dendrometer band data.](tables_figures/growth_curves_all.png){ width=100% }

@@

![**Extended Data Figure 3 | Landscapes of relationships between the day of year on which 25% of annual growth is achieved ($DOY_{25}$) and temperature in prior weeks for ring- and diffuse-porous trees at the Smoithsonian Conservation Biology Institute (SCBI) and Harvard Forest.** Shown are matrices of $\beta$ coefficients from first-order linear regressions between mean maximum temperature ($T_{max}$) and $DOY_{25}$. Window Open and Window Close indicate number of weeks prior to $DOY_{25}$ (listed in `r EDT_phenology_parameters`). Yellow shading indicates neutral relationships, while orange or red shading indicates that $DOY_{25}$ advances with increased $T_{max}$ over the given time window (negative $\beta$). Black circles indicate the critical temperature window selected based on minimization of $\Delta AICc$, the difference in Akaike Information Criterion corrected for small sample size relative to a null model. Critical temperature windows are listed in `r EDT_phenology_parameters`.](tables_figures/climwin_figure.png){ width=70% }

@@

![**Extended Data Figure 4 | Response of the timing of stem growth phenology to mean maximum temperatures during the critial temperature window (CTW).** Relationships between critical temperature window (CTW) $T_max$ and 25%, 50%, and 75% total stem growth (first row), the length of the peak growing season ($L_pgs$; second row), maximum growth rate ($g_max$; third row), and total seasonal radial stem growth ($\Delta DBH$; fourth row) are shown. Posterior predictions of each variable that did not include zero are represented with solid lines, while those that do include zero use dotted lines. 95% credible intervals are represented by bands. The days of year on which 25%, 50% and 75% of annual growth were acheived ($DOY_{25}$, $DOY_{50}$, and $DOY_{75}$, respectively) declined significantly with mean $T_{max}$ during their respective CTW. For each CTW $T_{max}$. Modeled values from dendrometer band measurements are represented by dots.](tables_figures/pheno_Tsensitivity_combo_patchwork_AIC.png){ width=90% }

@@

![**Extended Data Figure 5 | Relationship between growth parameters and mean maximum temperatures in April.** Relationships between critical temperature window (CTW) $T_max$ and 25%, 50%, and 75% total stem growth (first row), the length of the peak growing season ($L_pgs$; second row), maximum growth rate ($g_max$; third row), and total seasonal radial stem growth ($\Delta DBH$; fourth row) are shown. Posterior predictions of each variable that did not include zero are represented with solid lines, while those that do include zero use dotted lines. 95% credible intervals are represented by bands.](tables_figures/pheno_Tsensitivity_combo_patchwork_april.png){ width=90% }

>>>>

![**Extended Data Figure 2 | Phenological patterns of forest canopy greenness (top row) and stem growth of ring- and diffuse-porous trees, represented as both relative and cumulative fractions of total annual growth (middle and bottom rows, respectively), at the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest.** In the top row, canopy greenness is characterized using the two band Enhanced Vegetation Index (EVI2), with each line representing a year between 2000 and 2018. For stem growth, each line represents the growth of one tree over one year, as modeled based on a five-parameter logistic growth model to dendrometer band data.](tables_figures/growth_curves_all.png){ width=100% }

@@

![**Extended Data Figure 3 | Landscapes of relationships between the day of year on which 25% of annual growth is achieved ($DOY_{25}$) and temperature in prior weeks for ring- and diffuse-porous trees at the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest.** Shown are matrices of $\beta$ coefficients from first-order linear regressions between mean maximum temperature ($T_{max}$) and $DOY_{25}$. Window Open and Window Close indicate number of weeks prior to $DOY_{25}$ (listed in `r EDT_phenology_parameters`). Yellow shading indicates neutral relationships, while orange or red shading indicates that $DOY_{25}$ advances with increased $T_{max}$ over the given time window (negative $\beta$). Black circles indicate the critical temperature window selected based on minimization of $\Delta AICc$, the difference in Akaike Information Criterion corrected for small sample size relative to a null model. Critical temperature windows are listed in `r EDT_phenology_parameters`.](tables_figures/climwin_figure.png){ width=70% }

@@

![**Extended Data Figure 4 | Response of stem growth phenology and rates to mean maximum temperatures ($T_{max}$) during the spring critial temperature window (CTW) for ring- and diffuse-porous speces at the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest.** CTW was defined as the period over which $T_{max}$ was most strongly correlated with the day of year on which 25% of annual growth was achieved ($DOY_{25}$; `r EDT_phenology_parameters`, `r EDF_climwin`). Shown are relationships between mean $T_max$ over the CTW and days of the year on which 25%, 50%, and 75% total stem growth were achieved ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$, respectively; first row); the length of the peak growing season ($L_{pgs}$; second row); maximum growth rate ($g_{max}$; third row); and total seasonal radial stem growth ($\Delta DBH$; fourth row). Posterior predictions of each variable that did not include zero are represented with solid lines, while those that do include zero use dotted lines. 95% credible intervals are represented by bands. For both species groups at both sites, $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$ all declined significantly with mean $T_{max}$ during their respective CTW. Dots represent growth parameter values for individual tree-year combinations, which were derived by fitting a five-parameter logistic growth model to dendrometer band data.](tables_figures/pheno_Tsensitivity_combo_patchwork_AIC.png){ width=90% }

@@

![**Extended Data Figure 5 | Response of stem growth phenology and rates to April mean maximum temperature ($T_{max}$) for ring- and diffuse-porous speces at the Smithsonian Conservation Biology Institute (SCBI) and Harvard Forest.** Shown are relationships between April $T_max$ and days of the year on which 25%, 50%, and 75% total stem growth were achieved ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$, respectively; first row); the length of the peak growing season ($L_{pgs}$; second row); maximum growth rate ($g_{max}$; third row); and total seasonal radial stem growth ($\Delta DBH$; fourth row). Posterior predictions of each variable that did not include zero are represented with solid lines, while those that do include zero use dotted lines. 95% credible intervals are represented by bands. For both species groups at both sites, $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$ declined significantly with April $T_{max}$, with the exception of $DOY_{25}$ for ring-porous species at SCBI. Dots represent growth parameter values for individual tree-year combinations, which were derived by fitting a five-parameter logistic growth model to dendrometer band data.](tables_figures/pheno_Tsensitivity_combo_patchwork_april.png){ width=90% }
